### 2015-10-28 vmcai-2016-tex/evaluation.tex
Addressed reviewer concerns

\section{Evaluation}
\label{sec:evaluation}
%\begin{figure*}[t]
\begin{table*} [t]
  \centering
  \caption{Comparing \symtxt{} with the~\gsetxt{} and Lazier\# algorithms.}
  \scalebox{0.88}{\begin{tabular}{| c | c | r | r | r | r | r | r | r | r | r |}
  \hline
   \multirow{2}{*}{Method }&\multirow{2}{*}{ $k$ }
   &\multicolumn{3}{c|}{Time (seconds)} &\multicolumn{3}{c|}{States} &\multicolumn{3}{c|}{ Paths }\\
								&	&\gsetxt{} & L\#		&SL		&\gsetxt{}	& L\# & SL&\gsetxt{}	& L\# 	& SL\\
   \hline
    \multirow{3}{*}{LinkedList }			&3	& 0.91	& 1.21	& 0.69	& 2465	& 2844	& 99		&1656	& 1269	& 25\\
   		 						& 4	& 2.92	& 3.35	& 0.91	& 25774	& 29977	& 155	&17485	& 13550	& 39\\
   								& 5	& 20.78	& 19.47	& 1.59	& 341164	& 400296	& 223	&232743	& 181849	& 56\\
								& 6	& 280.56	& 299.19	& 2.36	&5447980	&6437201	& 303	&3731094	&2933027	& 76\\
								& 7	& -		& -		& 5.07	& -		&-		& 395	&-		&-		& 99\\
								& 8	& -		& -		& 17.49	& -		&-		& 499	&-		&-		& 125\\
								& 9	& -		& -		& 63.96	& -		&-		& 615	&-		&-		& 154\\
								& 10	& -		& -		& 206.93	& -		&-		& 743	&-		&-		& 186\\
    \hline
    \multirow{3}{*}{BinarySearchTree }	& 1	& 0.26	& 0.28	& 0.36	& 19		& 23		& 29		& 6		& 6		& 6\\
   		 						& 2	& 0.83	& 1.28	& 0.93	& 143	& 143	& 145	& 43		& 42		& 33\\
   								& 3	& 20.63	& 25.55	& 4.03	& 1953	& 1703	& 1485	& 515	& 515	& 328\\
								& 4	& -		& -		& 410.89	& -		&-		& 73635	&-		&-		& 15563\\
    \hline
      \multirow{3}{*}{TreeMap}			& 1	& 0.47	& 0.52	& 0.77	& 65		& 70		& 215	& 11		& 11		& 11\\
   		 						& 2	& 8.99	& 9.73	& 4.72	& 1009	& 942	& 3219	& 127	& 122	& 73\\
   								& 3	& -		& -		& 145.56	& -		& -		& 78695	& -		& -		& 887\\

@@

%\end{figure*}

%\subsection{Analysis}
The \symtxt{} algorithm is implemented as an extension to the Symbolic
PathFinder (SPF)~\cite{DBLP:journals/ase/PasareanuVBGMR13}
framework. In addition to the operations presented in this paper, the
implementation contains support for operations over integers,
calculating per-path preconditions and postconditions, as well as
generating test input heaps that exercise all feasible control flow
paths. Future work is adding support for floating point
operations, arrays, and bit-operations.

The \symtxt{} implementation uses \texttt{jConstraints} with the z3 solver \cite{ase2014-ghilrr,jpf2014-dghirr,deMouraBjorner08Z3}. The implementation takes advantage of incremental solving and employs caching for performance since only small portions of the heap constraint changes during the search.

SPF includes an implementation of~\gsetxt{} with lazy initialization.
In recent work, we implemented the Lazier and Lazier\# algorithms in
SPF~\cite{Hillery:2014}; these constitute the state of art approaches
case-splitting based lazy initialization techniques. The goal of our
experiment is to evaluate the efficacy of our approach with respect to
other state-of-the-art techniques for symbolic execution of programs
with unbounded complex data input. We empirically evaluate the
\symtxt{} algorithm by considering the following research question:
How does the cost of the \symtxt{} algorithm (SL) compare with that of
\gsetxt{} (GSE), Lazier, Lazier\#(L\#) algorithms?

The \emph{independent} variable in our study is the $k$-bound;
$k$-bounding bounds the length of a reference chain from the root of
the heap~\cite{Kiasan06}.  We select three dependent variables and
measures for our study: (i) time, (ii) states explored, and (iii)
paths generated. The \emph{time} is the total time taken by each
algorithm to explore the symbolic execution tree as well as total
constraint solving time. The \emph{states explored} represents the
number of nodes in the symbolic execution tree, and the \emph{paths
  generated} represents the number of unique paths in the symbolic
execution tree.

The data structures evaluated are a standard set commonly used in
analyses involving heap-manipulating
programs~\cite{Deng:2006,Deng:2007,boyapati2002korat,Ferrara:2014,Rosner:2015},
including a linked list, binary search tree, and red/black tree. We
use a repOk() method (a class invariant) for data structures in
object-oriented code~\cite{boyapati2002korat}. The repOk() methods
provide a mechanism to generate valid inputs for the methods under
test; symbolic execution has been used to generate test inputs using
repOk() methods in~\cite{visser2004test}. In this work, we define and
use repOk() methods to generate valid heaps with the appropriate
$k$-bounds before executing the methods under test. Note that this
allows us to have precision in checking properties of the heap that is
not possible in static analysis based techniques~\cite{Dillig:2011}.

The results of the experiments are presented in
Table~\ref{tab:results}.  Each row reports the results for the
specified $k$-bound for each artifact evaluated. The columns show the
total time in seconds, states explored, and paths generated for each
algorithm. The headings \gsetxt{}, L\#, and SL correspond to the GSE,
Lazier\#, and \symtxt{} algorithms, respectively. The results for the
Lazier algorithm are omitted because they are nearly identical to the
Lazier\# results for these artifacts. A table entry of '-' indicates
the analysis exceeded the allotted time bound of 1 hour.

The number of possible non-isomorphic heap configurations grows
exponentially for case-splitting techniques with a monotonic increase
in $k$, resulting a corresponding exponential increase in analysis
times . This is event in all the examples in
Table~\ref{tab:results}. The GSE and L\# algorithms are unable to
finish exploration in a time bound of one hour for $k\geq 7$ for the
LinkedList, $k\geq 4$ for the BinarySearchTree, and $k\geq 3$ for the
TreeMap examples. The improvement in analysis time for the \symtxt{}
over the state of the art case splitting techniques range from $4.8x$
for BinarySearchTree at $k$=3, to $118x$ for LinkedList at $k$=6. In
fact, for some $k$-bounds a number of experiments complete exploration
using the \symtxt{} algorithm in a few seconds whereas \gsetxt{} or
Lazier\# are unable to finish, e.g., BinarySearchTree for $k$=4 and
LinkedList for $k$=8.

The number of path explored by \symtxt{} algorithm are strictly less
than or equal to the number of paths explored by \gsetxt{} for all the
evaluated artifacts. We can use this result to do more efficient test
case generation. From a generated path and corresponding symbolic heap
we can use the solutions provided by the constraint solver to
instantiate a set of concrete heaps to provide as test input. These
provide a smaller test suite to achieve control-flow path coverage as
compared to GSE and Lazier\#. Note that the ability to perform test
input generation is another advantage of \symtxt{} technique over
static analysis approaches.

The number of states varies between algorithms, for example, \gsetxt{}
has additional points of nondeterminism during field reads, but in
contrast reference compares are completely deterministic. Thus, in
example programs with large numbers of reference compares, such as
TreeMap, state counts for Lazier\# and the \symtxt{} algorithm may
exceed those for \gsetxt{}. We observed that the additional states
generated by the summary heap algorithm are unsatisfiable at the point
of reference compares; this is why they do not contribute any
additional branches in the final symbolic execution tree. Furthermore,
the larger state count and the corresponding satisfiability checks on
the constraint solver does not increase the overall runtime of the
technique.

In summary, the benefits of avoiding case-splitting based
non-determinism outweigh the increased complexity in the constraints
over heaps due to the advances made in SMT solvers in these
examples. In fact, the \symtxt{} algorithm can analyze certain types
of programs with orders of magnitude greater efficiency than with
\gsetxt{} or Lazier\#, while covering exactly the same feasible
control flow paths in the program.

>>>>

\section{Evaluation}
\label{sec:evaluation}
%\begin{figure*}[t]
%\end{figure*}

%\subsection{Analysis}
The \symtxt{} algorithm is implemented as an extension to the Symbolic
PathFinder (SPF)
framework~\cite{DBLP:journals/ase/PasareanuVBGMR13}. In addition to the operations presented in this paper, the
implementation contains support for operations over integers,
calculating per-path preconditions and postconditions, as well as
generating test input heaps that exercise all feasible control flow
paths. Future work is adding support for floating point
operations, arrays, and bit-operations.

The \symtxt{} implementation uses \texttt{jConstraints} with the z3 solver \cite{ase2014-ghilrr,jpf2014-dghirr,deMouraBjorner08Z3}. The implementation takes advantage of incremental solving and employs caching for performance since only small portions of the heap constraint change during the search.

SPF includes an implementation of~\gsetxt{} with lazy initialization.
In recent work, we implemented the Lazier and Lazier\# algorithms in
SPF~\cite{Hillery:2014}; these constitute the state of art approaches
to case-splitting based lazy initialization techniques. The goal of our
experiment is to evaluate the efficacy of our approach with respect to these
other techniques for symbolic execution of programs
with unbounded complex data input. The empirical study tries to answer the following research question:
How does the cost of the \symtxt{} algorithm compare with that of the \gsetxt{} and Lazier\# algorithms?

The \emph{independent} variable in the study is the $k$-bound;
$k$-bounding bounds the length of a reference chain from the root of
the heap~\cite{Deng:2006}.  The study selects three dependent variables and
measures: (i) time, (ii) states explored, and (iii)
paths generated. The \emph{time} is the total time taken by each
algorithm to explore the symbolic execution tree as well as total
constraint solving time. The \emph{states explored} represents the
number of nodes in the symbolic execution tree, and the \emph{paths
  generated} represents the number of unique paths in the symbolic
execution tree.

The data structures evaluated are a standard set that is commonly used in
analyses involving heap-manipulating
programs~\cite{Deng:2006,Deng:2007,boyapati2002korat,Ferrara:2014,Rosner:2015},
including a linked list, binary search tree, and red/black tree. The actual tests
use a repOk() method (a class invariant) for data structures in
object-oriented code to generate valid inputs for the methods under
test via symbolic execution~\cite{boyapati2002korat,visser2004test}. Note that this
allows us to have precision in checking properties of the heap that is
not possible in static analysis based techniques (i.e., the symbolic execution generated heaps satisfy the repOK() invariant)~\cite{Dillig:2011}.

\begin{table*} [t]
  \centering
  \caption{Comparing \symtxt{} with the~\gsetxt{} and Lazier\# algorithms.}
  \scalebox{0.88}{\begin{tabular}{| c | c | r | r | r | r | r | r | r | r | r |}
  \hline
   \multirow{2}{*}{Method }&\multirow{2}{*}{ $k$ }
   &\multicolumn{3}{c|}{Time (seconds)} &\multicolumn{3}{c|}{States} &\multicolumn{3}{c|}{ Paths }\\
								&	&\gsetxt{} & L\#		&SL		&\gsetxt{}	& L\# & SL&\gsetxt{}	& L\# 	& SL\\
   \hline
    \multirow{3}{*}{LinkedList }			&3	& 0.91	& 1.21	& 0.69	& 2465	& 2844	& 99		&1656	& 1269	& 25\\
   		 						& 4	& 2.92	& 3.35	& 0.91	& 25774	& 29977	& 155	&17485	& 13550	& 39\\
   								& 5	& 20.78	& 19.47	& 1.59	& 341164	& 400296	& 223	&232743	& 181849	& 56\\
								& 6	& 280.56	& 299.19	& 2.36	&5447980	&6437201	& 303	&3731094	&2933027	& 76\\
								& 7	& -		& -		& 5.07	& -		&-		& 395	&-		&-		& 99\\
								& 8	& -		& -		& 17.49	& -		&-		& 499	&-		&-		& 125\\
								& 9	& -		& -		& 63.96	& -		&-		& 615	&-		&-		& 154\\
								& 10	& -		& -		& 206.93	& -		&-		& 743	&-		&-		& 186\\
    \hline
    \multirow{3}{*}{BinarySearchTree }	& 1	& 0.26	& 0.28	& 0.36	& 19		& 23		& 29		& 6		& 6		& 6\\
   		 						& 2	& 0.83	& 1.28	& 0.93	& 143	& 143	& 145	& 43		& 42		& 33\\
   								& 3	& 20.63	& 25.55	& 4.03	& 1953	& 1703	& 1485	& 515	& 515	& 328\\
								& 4	& -		& -		& 410.89	& -		&-		& 73635	&-		&-		& 15563\\
    \hline
      \multirow{3}{*}{TreeMap}			& 1	& 0.47	& 0.52	& 0.77	& 65		& 70		& 215	& 11		& 11		& 11\\
   		 						& 2	& 8.99	& 9.73	& 4.72	& 1009	& 942	& 3219	& 127	& 122	& 73\\
   								& 3	& -		& -		& 145.56	& -		& -		& 78695	& -		& -		& 887\\

@@

The results of the experiments are presented in
Table~\ref{tab:results}.  Each row reports the results for the
specified $k$-bound for each artifact evaluated. The columns show the
total time in seconds, states explored, and paths generated for each
algorithm. The headings \gsetxt{}, L\#, and SL correspond to the GSE,
Lazier\#, and \symtxt{} algorithms, respectively. A table entry of '-' indicates
the analysis exceeded the allotted time bound of 1 hour.

The number of possible non-isomorphic heap configurations grows
exponentially for case-splitting techniques with a monotonic increase
in $k$, resulting in a corresponding exponential increase in analysis
times. This is evident in all the examples in
Table~\ref{tab:results}. The GSE and L\# algorithms are unable to
finish exploration in a time bound of one hour for $k\geq 7$ for the
LinkedList, $k\geq 4$ for the BinarySearchTree, and $k\geq 3$ for the
TreeMap examples. The improvement in analysis time for the \symtxt{}
over the state of the art case splitting techniques range from $4.8x$
for BinarySearchTree at $k$=3, to $118x$ for LinkedList at $k$=6. In
fact, for some $k$-bounds, a number of experiments complete exploration
using the \symtxt{} algorithm in a few seconds whereas \gsetxt{} or
Lazier\# are unable to finish, e.g., BinarySearchTree for $k$=4 and
LinkedList for $k$=8.

The number of path explored by the \symtxt{} algorithm are strictly less
than or equal to the number of paths explored by \gsetxt{} for all the
evaluated artifacts. This result means it is possible to do more efficient test
case generation by using the generated path and corresponding symbolic heap
solutions provided by the constraint solver to
instantiate a set of concrete heaps. These heaps
provide a smaller test suite to achieve control-flow path coverage as
compared to GSE and Lazier\#. Note that the ability to perform test
input generation is another advantage of the \symtxt{} technique over
static analysis approaches.

The number of states varies between algorithms, for example, \gsetxt{}
has additional points of nondeterminism during field reads, but in
contrast reference compares are completely deterministic. Thus, in
example programs with large numbers of reference compares, such as
TreeMap, state counts for Lazier\# and the \symtxt{} algorithm may
exceed those for \gsetxt{}. Observe that the additional states
generated by the summary heap algorithm are unsatisfiable at the point
of reference compares; this is why they do not contribute any
additional branches in the final symbolic execution tree. More critically,
the larger state count and the corresponding satisfiability checks on
the constraint solver does not increase the overall runtime of the
technique.

In summary, the benefits of avoiding case-splitting based
non-determinism outweigh the increased complexity in the constraints
over heaps due to the advances made in SMT solvers in these
examples. In fact, the \symtxt{} algorithm can analyze certain types
of programs with orders of magnitude greater efficiency than that of
\gsetxt{} or Lazier\#, while covering exactly the same feasible
control flow paths in the program.

### 2015-10-28 vmcai-2016-tex/lazy.tex
Addressed reviewer concerns

\section{GSE}
\label{app:gse}
This section introduces \gsetxt\ semantics on which \symtxt\ builds. GSE and its variants have the same non-deterministic choice given a reference to an uninitialized object; that reference can point to null, a new instance of an object with the correct type, or to an object of the same type that has been instantiated previously by GSE \cite{GSE03,KiasanKunit,Cadar:2008,Rosner:2015}. The objects instantiated for new instances are referred to as the \emph{input heap}. Only these objects comprise the potential alias set when GSE encounters references to uninitialized objects. In general, the GSE search space branches at uninitialized objects, and the number of branches depends on the size of the input heap. The function
$\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) =
\{\lp\phi\ \cfgnt{l}\rp\ ...\}$ returns constraint-location
pairs where the field $\cfgnt{f}$ is uninitialized.

\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]

The lazy initialization rules for GSE on the symbolic heap are in
\figref{fig:lazyInit}: NULL, new, and alias. The rules rely on fresh references and locations that strictly increase so it is possible to minimize over a set to find the first created. Further, references are partitioned to support latter proofs: $\mathrm{init}_\cfgnt{r}()$ for the input
  heap
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
    literals}.
  In general, as shown in the next section,
  only input heap references appear in constraints to express potential aliasing, and only stack references appear in environments, expressions, or continuations.  Finally, the \emph{isInit} function is true for initialized references from the input heap (i.e., potential aliases).

How the lazy
initialization is used is defined in \figref{fig:lazy} with the $\rgse$ relation collecting all the rules into a single relation on states. This relation is used later by the proof. GSE initialization takes place on the field-access rule in \figref{fig:lazy},
using the $\rightarrow_I^*$ relation from \figref{fig:lazyInit}, to ensure the accessed field is
instantiated. Initialization in GSE never
happens for more than one object on any use of $\rightarrow_I^*$:
the set $\Lambda$ is either empty or contains exactly one location. This property is an artifact of how GSE case splits when it instantiates: each choice is a new unique heap. This changes in the next section with the new symbolic initialization that collects all the choices into a single heap using guarded value sets.

>>>>

\section{GSE}
\label{app:gse}
This section introduces \gsetxt\ semantics on which \symtxt\ builds. GSE and its variants have the same non-deterministic choice given a reference to an uninitialized object; that reference can point to null, a new instance of an object with the correct type, or to an object of the same type that has been instantiated previously by GSE \cite{GSE03,KiasanKunit,Cadar:2008,Rosner:2015}. The objects instantiated for new instances are referred to as the \emph{input heap}. Only these objects comprise the potential alias set when GSE encounters references to uninitialized objects. In general, the GSE search space branches at uninitialized objects, and the number of branches depends on the size of the input heap. The function

The lazy initialization rules for GSE on the symbolic heap are in
\figref{fig:lazyInit}: NULL, new, and alias.  The symbol $\cfgnt{C}$ represents a type (or class) while $\mathit{fields}\lp \mathrm{C}\rp$ returns the fields in the type. The function $\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) =
\{\lp\phi\ \cfgnt{l}\rp\ ...\}$ returns constraint-location
pairs where the field $\cfgnt{f}$ is uninitialized.
  \[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]
The function $\mathbb{S}\lp \phi \wedge \phi^\prime \rp$ returns true
if there is a satisfying assignment of references for $\phi \wedge
\phi^\prime$; otherwise it returns false.

The rules rely on fresh references and locations that strictly increase so it is possible to minimize over a set to find the first created (i.e., references and locations are ordered). As such, the $\mathrm{min}_l$ function is able to return $(\phi_x\ \cfgnt{l}_x)$ the earliest created uninitialized location in a set, and similarly, $\mathrm{min}_r$ is able to return the earliest created reference in a set. Further, references are partitioned to support latter proofs: $\mathrm{init}_\cfgnt{r}()$ for the input
  heap
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
    literals}.
  In general, as shown in the next section,
  only input heap references appear in constraints to express potential aliasing, and only stack references appear in environments, expressions, or continuations.  Finally, the \emph{isInit} function is true for initialized references from the input heap (i.e., potential aliases).

How the lazy
initialization is used is defined in \figref{fig:lazy} with the $\rgse$ relation collecting all the rules into a single relation on states.  GSE initialization takes place on the field-access rule in \figref{fig:lazy},
using the $\rightarrow_I^*$ relation from \figref{fig:lazyInit}, to ensure the accessed field is
instantiated. Initialization in GSE never
happens for more than one object on any use of $\rightarrow_I^*$:
the set $\Lambda$ is either empty or contains exactly one location. This property is an artifact of how GSE case splits when it instantiates: each choice, NULL, new, or an alias, is a new unique heap. This changes in the next section with the new symbolic initialization that collects all the choices into a single heap using guarded value sets. The field-write rule also uses $\theta$ to represent a set of constraint-location pairs, which in GSE, again should always be a singleton set for the same reason as previously mentioned.

### 2015-10-28 vmcai-2016-tex/rw.tex
Addressed reviewer concerns

%\subsection{Other Heap Representations in Symbolic Execution}
%Need to talk about GSE, DSE, SAGE, PEX
%GSE
Symbolic initialization also draws inspiration from lazy initialization, the core idea of generalized symbolic execution (\gsetxt{})\cite{GSE03}. Several projects have used lazy initialization to conduct symbolic execution on programs with more general types,
including references and
arrays~\cite{KiasanKunit,Cadar:2008,Deng:2007,Rosner:2015,Blackshear:2013,Filieri:2015}. However, \gsetxt{} techniques branch over
multiple copies of the system state during dereferencing operations,
exacerbating path explosion.

@@

%DSE
A number of dynamic symbolic execution (\dsetxt{}) methods use some form 
of symbolic heap representation, including PEX~\cite{Tillmann:2008} and SAGE~\cite{Elkarablieh:2009}. These~\dsetxt{} methods have shown a high degree of utility in practical applications with a wide variety of real-world programs. PEX and SAGE use array theories to represent program state instead of guarded value sets.
SAGE has been proven complete for programs whose memory allocations are independent of their inputs. This work is complementary to PEX and SAGE, and we believe could be leveraged to further improve those tools.

Several separation logic solvers have been proven to be sound and complete for heaps with linked lists~\cite{Navarro:2011,Cook:2011,Berdine:2005}, trees~\cite{Piskac:2014}, or data structures satisfying user-supplied invariants~\cite{Brotherston:2014}. Separation logic solvers can also reason about entailment which is a subject of future work for symbolic initialization. Solutions in separation logic do not lend themselves to test case generation is  goal in this work.

>>>>

%\subsection{Other Heap Representations in Symbolic Execution}
%Need to talk about GSE, DSE, SAGE, PEX
%GSE
Symbolic initialization also draws inspiration from lazy initialization, the core idea of \gsetxt{}\cite{GSE03}. Several projects have used lazy initialization to conduct symbolic execution on programs with more general types,
including references and
arrays~\cite{KiasanKunit,Cadar:2008,Blackshear:2013,Filieri:2015}. Improvements to the basic lazy initialization algorithm have been proposed, including delaying aliasing choices~\cite{Deng:2007}, or checking initializations against invariants as they occur~\cite{Braione:2015,Rosner:2015}. However, these \gsetxt{} techniques branch over
multiple copies of the system state during dereferencing operations,
exacerbating path explosion.

@@

%DSE
A number of dynamic symbolic execution (\dsetxt{}) methods use some form 
of symbolic heap representation, including PEX~\cite{Tillmann:2008} and SAGE~\cite{Elkarablieh:2009}. These~\dsetxt{} methods have shown a high degree of utility in practical applications with a wide variety of real-world programs. PEX and SAGE use array theories to represent program state instead of guarded value sets.
SAGE has been proven complete for programs whose memory allocations are independent of their inputs. This work is complementary to PEX and SAGE and could be leveraged to further improve those tools.

Several separation logic solvers have been proven to be sound and complete for heaps with linked lists~\cite{Navarro:2011,Cook:2011,Berdine:2005}, trees~\cite{Piskac:2014}, or data structures satisfying user-supplied invariants~\cite{Brotherston:2014}. Separation logic solvers can also reason about entailment which is a subject of future work for symbolic initialization. Solutions in separation logic do not lend themselves to test case generation which is a contribution of the new approach in this paper.
